---
title: "Computational Musicology Portfolio"
subtitle: "K-pop vs Cantopop (Current Available Features)"
format:
  dashboard:
    orientation: columns
    scrolling: true
execute:
  echo: false
  warning: false
  message: false
---

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)

normalize_names <- function(df) {
  names(df) <- names(df) |>
    tolower() |>
    gsub("[^a-z0-9]+", "_", x = .) |>
    gsub("^_|_$", "", x = .)
  df
}

kpop <- read_csv("kpop_2020s.csv", show_col_types = FALSE) |> normalize_names() |> mutate(collection = "K-pop 2020s")
cantopop <- read_csv("cantopop_2020s.csv", show_col_types = FALSE) |> normalize_names() |> mutate(collection = "Cantopop 2020s")

tracks <- bind_rows(kpop, cantopop) |>
  mutate(
    release_date = as.Date(release_date),
    duration_min = duration_ms / 60000
  ) |>
  rename(artist_names = artist_name_s)

numeric_features <- c("danceability", "energy", "valence", "tempo", "loudness", "acousticness", "speechiness", "liveness", "duration_min", "popularity")

feature_long <- tracks |>
  select(collection, all_of(numeric_features)) |>
  pivot_longer(-collection, names_to = "feature", values_to = "value")

top_artists <- tracks |>
  count(collection, artist_names, sort = TRUE) |>
  group_by(collection) |>
  slice_head(n = 10) |>
  ungroup()
```

## Overview

### Key stats

```{r}
track_n <- nrow(tracks)
artist_n <- tracks |> distinct(artist_names) |> nrow()
avg_pop <- mean(tracks$popularity, na.rm = TRUE)

summary_tbl <- tibble(
  Metric = c("Total tracks", "Unique artists", "Average popularity"),
  Value = c(track_n, artist_n, number(avg_pop, accuracy = 0.1))
)

knitr::kable(summary_tbl)
```

### Tracks by release year

```{r}
tracks |>
  mutate(release_year = format(release_date, "%Y")) |>
  count(collection, release_year) |>
  ggplot(aes(x = release_year, y = n, fill = collection)) +
  geom_col(position = "dodge") +
  labs(x = "Release year", y = "Track count", fill = "Collection") +
  theme_minimal()
```

## Genre Comparison

### Feature distribution

```{r}
feature_long |>
  filter(feature %in% c("danceability", "energy", "valence", "tempo")) |>
  ggplot(aes(x = collection, y = value, fill = collection)) +
  geom_boxplot(alpha = 0.8, outlier.alpha = 0.4) +
  facet_wrap(~ feature, scales = "free_y") +
  labs(x = "", y = "Value") +
  theme_minimal() +
  theme(legend.position = "none")
```

### Energy vs Valence

```{r}
tracks |>
  ggplot(aes(x = energy, y = valence, color = collection)) +
  geom_point(alpha = 0.7, size = 2) +
  labs(x = "Energy", y = "Valence", color = "Collection") +
  theme_minimal()
```

### Loudness vs Acousticness

```{r}
tracks |>
  ggplot(aes(x = acousticness, y = loudness, color = collection)) +
  geom_point(alpha = 0.7, size = 2) +
  labs(x = "Acousticness", y = "Loudness (dB)", color = "Collection") +
  theme_minimal()
```

## Artists and Tracks

### Top artists by track count

```{r}
top_artists |>
  ggplot(aes(x = reorder(artist_names, n), y = n, fill = collection)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ collection, scales = "free_y") +
  coord_flip() +
  labs(x = "Artist", y = "Track count") +
  theme_minimal()
```

### Track table (current data)

```{r}
tracks |>
  select(collection, track_name, artist_names, release_date, popularity, danceability, energy, valence, tempo, loudness) |>
  arrange(collection, desc(popularity)) |>
  slice_head(n = 30) |>
  knitr::kable()
```

## Placeholders for Future Features

### Chroma / Pitch Features (to add later)

::: {.callout-note}
This section is intentionally a placeholder. Add chroma/pitch-level features when available, then include pitch class heatmaps and key-profile comparisons.
:::

### Timbre Features (to add later)

::: {.callout-note}
This section is intentionally a placeholder. Add timbre vectors (e.g., MFCC-based descriptors) and include clustering/PCA visualizations.
:::

### Section-Level Dynamics (to add later)

::: {.callout-note}
This section is intentionally a placeholder. Add section/segment-level data for temporal loudness and structural form analysis.
:::
